---
title: "Assignment 5"
author: "Jesse Y"
date: "2025-03-15"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}

# Load required libraries
library(tidyverse)
library(clarify)
library(broom)
library(texreg)
library(lmtest)
library(rmdformats)

# Suppress warnings/messages for cleaner output
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# **Introduction**

This assignment extends Assignment 4, using post-estimation methods via clarify to enhance interpretation of our logistic regression model predicting transmission type (am) in the mtcars dataset.

We will:

- Simulate model coefficients to assess uncertainty.

- Compute Average Marginal Effects (AME) to quantify the effects of predictors.

- Estimate Predicted Probabilities at different mpg values for deeper insight.

- Evaluate Model Fit using Likelihood Ratio Test, AIC, and BIC.


# **Data Preparation**

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# Load mtcars dataset
mtcars_data <- mtcars

# Convert 'am' (Transmission Type) into a binary factor
mtcars_data <- mtcars_data %>%
  mutate(am = factor(am, labels = c("Automatic", "Manual")),
         hp_group = ifelse(hp > median(hp), "High", "Low"))  # Dichotomize horsepower

# Check dataset structure
str(mtcars_data)

```


# **Logistic Regression Model**

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# Model: Predicting Transmission Type
m1 <- glm(am ~ mpg + hp_group + wt, data = mtcars_data, family = binomial)
summary(m1)
```

Key Findings:

- mpg is not statistically significant (p = 0.6277), meaning fuel efficiency does not strongly predict transmission type. 

- However, weight (wt) is statistically significant (p = 0.0138), indicating that heavier cars are more likely automatic.  

- hp_group is not statistically significant, suggesting that horsepower (high vs. low) alone does not explain transmission type.


# **Post-Estimation with clarify**

## 1. Simulating Model Coefficients

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
set.seed(123)
sim_coefs <- sim(m1, n = 1000)
summary(sim_coefs)
```
Significance of This Approach:

- Simulations allow us to estimate uncertainty around our model coefficients.

- This avoids over-reliance on single coefficient values.

## 2. Average Marginal Effects (AME)

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
sim_ame_results <- sim_ame(sim_coefs, var = "mpg", contrast = "rd", verbose = FALSE)
summary(sim_ame_results)

# Plot AME
plot(sim_ame_results, main = "Average Marginal Effect of MPG on Transmission")
```

Findings: Average Marginal Effects (AME)

- The AME of mpg (-0.0116) suggests that increasing mpg has an extremely small effect on the probability of a car having a manual transmission.  

- This confirms that mpg is not statistically significant, and transmission choice is likely influenced by other variables such as weight.

## 3. Predictions at Set Values

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# Define mpg values for prediction
sim_pred <- sim_setx(sim_coefs, x = list(mpg = c(15, 25)))  
summary(sim_pred)

# Plot predicted probabilities
plot(sim_pred, main = "Predicted Probability of Transmission Type at Different MPG Values")
```


Findings: Predicted Probabilities

- At 15 mpg, the probability of manual transmission is 18%.

- At 25 mpg, the probability drops to 4.4%, meaning cars with higher mpg are actually less likely to have a manual transmission in this dataset.

- This suggests that other factors (like weight) may have a stronger influence on transmission choice than mpg alone.


# **Model Selection and Fit**

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
lrtest(m1)   # Likelihood Ratio Test
AIC(m1)      # Akaike Information Criterion
BIC(m1)      # Bayesian Information Criterion
```

Findings: Model Selection and Fit

- The Likelihood Ratio Test (p < 0.001) confirms that our model is statistically significant, meaning the predictors (mpg, hp_group, wt) help explain transmission type.  

- Lower AIC/BIC values suggest our model strikes a balance between fit and complexity.


# Key Findings

1. MPG has a small, non-significant effect on transmission type.

2. Weight (wt) is a strong predictor - heavier cars are more likely automatic.

3. Clarify confirms these findings:

- Marginal effects plot shows little impact from mpg.

- Predicted probabilities demonstrate unexpected trends in mpg effects.


# Conclusion

This assignment demonstrated how clarify enhances logistic regression analysis by:

- Estimating uncertainty through simulations.

- Providing meaningful marginal effects instead of raw log-odds.

- Visualizing predicted probabilities to communicate findings effectively.
